---
title: "Data analysis"
author: "Xueying Feng"
date: "4/11/2020"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

getwd()
setwd("~/Desktop/Environ 872/ENV872 Project_XY_Feng")

# Load your packages
library(plyr) #Tools for Splitting, Applying and Combining Data
library(tidyverse) #Getting data frames to tidy
library(lubridate) #For dates and date-times

require(ggplot2)
#install.packages("ggrepel") 
require(ggrepel)

#install.packages("usmap")
library(usmap)

# Set your ggplot theme
mytheme <- theme_minimal(base_size = 12, base_family = "Times") + 
  theme(axis.text.x = element_text(color = "DarkGrey"),
        legend.position = "top") 
theme_set(mytheme)

# Load your datasets
Historical_Global_processed <- read.csv("./Data/Processed/Historical_Global_processed.csv")
Historical_China_processed <- read.csv("./Data/Processed/Historical_China_processed.csv")
Covid19_Global_processed <- read.csv("./Data/Processed/Covid19_Global_processed.csv")
Covid19_China_processed <- read.csv("./Data/Processed/Covid19_China_processed.csv")
Covid19_US_processed <- read.csv("./Data/Processed/Covid19_US_processed.csv")

```

## Anlysis trend of confirmed cases in China

### Trend of Total confirmed cases in China

```{r Select China data}
# Selsct data from China
str(Historical_Global_processed)
Historical_China = Historical_Global_processed[Historical_Global_processed$country == 'China',]
head(Historical_China)
```


```{r China_Total_Trend.plot, fig.cap=paste("Trend of confirmed case in China")}
# Create a ggplot depicting cases incereasing over time
China_Total_Trend.plot <- 
  ggplot(Historical_China, aes(x=time, y=cum_confirm)) +
  geom_point(colour = "#e6550d") +
  geom_line(colour = "#d95f0e") +
  labs(x = "Time", 
       y = "Confirmed Cases") + 
  scale_x_date(date_labels = "%Y-%m-%d")

print(China_Total_Trend.plot)
```


### Trend of confirmed cases in each province in China 

```{r Explore Historical_China_processed dataset, results='hide', message=FALSE, warning=FALSE}
# Check data
head(Historical_China_processed)
str(Historical_China_processed)

# Change date column to date
Historical_China_processed$time <- as.Date(Historical_China_processed$time, 
                                           format = "%m/%d/%y") 
head(Historical_China_processed)
class(Historical_China_processed$time)

```

### Trend of confirmed cases in Hubei, China

```{r Select all Hubei province data, results='hide', message=FALSE, warning=FALSE}
# Select Hubei province data
Historical_Hubei = Historical_China_processed[Historical_China_processed$province == 'Hubei',]
head(Historical_Hubei)

# Group by Hubei cities
Historical_HubeiCityV2 <- Historical_Hubei %>%
  select(time, city:cum_dead) 

# Save as csv
write.csv(Historical_HubeiCityV2,
          file = "./Data/Processed/Historical_HubeiCity.csv",row.names=FALSE)
```

```{r HubeiCity_Trend.plot, fig.cap=paste("Trend of each city in Hubei province, China")}
# Create a ggplot depicting cases incereasing over time
HubeiCity_Trend.plot <- ggplot(Historical_HubeiCityV2, 
                               aes(x=time, y=cum_confirm, color=city)) +
    geom_line(alpha = 0.95, size = 0.5) +
    geom_text_repel(aes(label=city),
                    function(Historical_HubeiCityV2) 
                      Historical_HubeiCityV2[Historical_HubeiCityV2$time == as.Date("2020-04-16"),])+
    mytheme +
    theme(legend.position = "none") +
    labs(x=expression(paste("Time"))) + 
    labs(y=expression(paste("Comfimed Cases")))+
    labs(color="city") +
    scale_x_date(date_labels = "%Y-%m-%d")

print(HubeiCity_Trend.plot)
```


### Trend of confirmed cases in Hubei cities, China (except Wuhan)

```{r Select Hubei cities (except Wuhan), results='hide', message=FALSE, warning=FALSE}
# Group by Hubei cities
Historical_HubeiCityV3 <- Historical_HubeiCityV2 %>%
  filter(city!= "Location TBD" & city!= "Wuhan") 

# Save as csv
write.csv(Historical_HubeiCityV3,
          file = "./Data/Processed/Historical_HubeiCity (except Wuhan).csv", row.names=FALSE)
```

```{r No_Wuhan.plot, fig.cap=paste("Trend of each city in Hubei province,China (except Wuhan)")}
# Create a ggplot depicting cases incereasing over time
No_Wuhan.plot <- ggplot(Historical_HubeiCityV3, aes(x=time, y=cum_confirm, color=city)) +
    geom_line(alpha = 0.95, size = 0.5) +
    geom_text_repel(aes(label=city),
                    function(Historical_HubeiCityV3) 
                      Historical_HubeiCityV3[Historical_HubeiCityV3$time == as.Date("2020-04-16"),])+
    mytheme +
    theme(legend.position = "none") +
    labs(x=expression(paste("Time"))) + 
    labs(y=expression(paste("Comfimed Cases")))+
    labs(color="city") +
    scale_x_date(date_labels = "%Y-%m-%d")

print(No_Wuhan.plot)
```


### Trend of confirmed cases in China provinces(except Hubei)

```{r Select provinces data (except Hubei) , results='hide', message=FALSE, warning=FALSE}
# Remove Hubei province data
Historical_ChinaProvince = Historical_China_processed[Historical_China_processed$province != 'Hubei',]
head(Historical_ChinaProvince)

Historical_ChinaProvinceV2 <- Historical_ChinaProvince %>%
  group_by(time,province) %>%
  summarise(total_confirm = sum(cum_confirm),
            total_heal = sum(cum_heal),
            total_dead = sum(cum_dead)) 

# Save as csv
write.csv(Historical_ChinaProvinceV2,
          file = "./Data/Processed/Historical_ChinaProvince (except Hubei).csv", row.names=FALSE)
```

```{r OtherProvinces.plot, fig.cap=paste("Trend of each province in China (except Hubei)")}
# Create a ggplot depicting cases incereasing over time
OtherProvinces.plot <- ggplot(Historical_ChinaProvinceV2, aes(x=time, y=total_confirm, color=province)) +
    geom_line(alpha = 0.95, size = 1) +
    geom_text_repel(aes(label=province),
                    function(Historical_ChinaProvinceV2) 
                      Historical_ChinaProvinceV2[Historical_ChinaProvinceV2$time == as.Date("2020-04-16"),])+
    mytheme +
    theme(legend.position = "none") +
    labs(x=expression(paste("Time"))) + 
    labs(y=expression(paste("Comfimed Cases"))) +
    labs(color="province") +
    scale_x_date(date_labels = "%Y-%m-%d")

print(OtherProvinces.plot)
```


### Trend of confirmed cases of top ten provinces in China (except Hubei)
```{r Select top ten provinces (except Hubei), results='hide', message=FALSE, warning=FALSE}
# Filter out the top ten provinces with the highest number of diagnoses (except Hubei)
TopTen_provinces <- Historical_ChinaProvinceV2 %>% 
  filter(time >= as.Date("2020-04-16")) %>%
  top_n(10, total_confirm) %>%
  arrange(desc(total_confirm)) 

TopTen_provinces <- pull(TopTen_provinces, province)

TenProvinces_China <- filter(Historical_ChinaProvinceV2, province %in% TopTen_provinces) %>%
  arrange(desc(total_confirm))

head(TenProvinces_China)

# Save as csv
write.csv(TenProvinces_China,
          file = "./Data/Processed/TenProvinces_China (except Hubei).csv", row.names=FALSE)
```

```{r TenProvinces_trend.plot, fig.cap=paste("Trend of top ten provinces, China (except Hubei)")}
# Draw plot
TenProvinces_trend.plot <- ggplot(TenProvinces_China, aes(x = time, y =total_confirm, color=province)) +
    geom_line(alpha = 0.95, size = 1) +
    geom_text_repel(aes(label=province),
                    function(Historical_ChinaProvinceV2) 
                      Historical_ChinaProvinceV2[Historical_ChinaProvinceV2$time == as.Date("2020-04-16"),])+
    mytheme +
    theme(legend.position = "none") +
    labs(x=expression(paste("Time"))) + 
    labs(y=expression(paste("Comfimed Cases"))) +
    labs(color="province") +
    scale_x_date(date_labels = "%Y-%m-%d")

print(TenProvinces_trend.plot)
```


## Anlysis Global trend of confirmed cases of top ten countries (except China) 

```{r Select top ten countries (except China) }
# Remove China data
Historical_Global = Historical_Global_processed[Historical_Global_processed$country != 'China',]
str(Historical_Global)

# Filter out the top ten countries with the highest number of diagnoses (except China)
TopTen_countries <- Historical_Global %>% 
  filter(time >= as.Date("2020-04-16")) %>%
  top_n(10, cum_confirm) %>%
  arrange(desc(cum_confirm)) 

# Selects a column in a data frame and transforms it into a vector
TopTen_countries <- pull(TopTen_countries, country)

Historical_TopTen <- filter(Historical_Global, country %in% TopTen_countries) %>%
  arrange(desc(cum_confirm))
head(Historical_TopTen)

# Save as csv
write.csv(Historical_TopTen,
          file = "./Data/Processed/TopTen_countries_trend(except China).csv", row.names=FALSE)
```

```{r TopTenCountries.plot, fig.cap=paste("Trend of Top Ten Countries (except China)")}
# Draw plot
TopTenCountries.plot <- ggplot(Historical_TopTen, aes(x = time, y =cum_confirm, color=country)) +
    geom_line(alpha = 0.95, size = 1) +
    geom_text_repel(aes(label=country),
                    function(Historical_TopTen) 
                      Historical_TopTen[Historical_TopTen$time == as.Date("2020-04-16"),])+
    mytheme +
    theme(legend.position = "none") +
    labs(x=expression(paste("Time"))) + 
    labs(y=expression(paste("Comfimed Cases"))) +
    labs(color="Country") +
    scale_x_date(date_labels = "%Y-%m-%d")

print(TopTenCountries.plot)
```


## Analysis trend of confirmed cases in United States. 

```{r Select United States data,results='hide', message=FALSE, warning=FALSE}
Historical_USTrend <- Historical_Global_processed[Historical_Global_processed$country == 'United States',] %>%
  na.omit(Historical_USTrend)

tail(Historical_USTrend)

# Save as csv
write.csv(Historical_USTrend,
          file = "./Data/Processed/Historical_USTrend.csv",row.names=FALSE)

```

### Trend of total confirmed case in United States.

```{r Wrangle United States data, results='hide', message=FALSE, warning=FALSE}
Historical_USTrend <- read.csv("./Data/Processed/Historical_USTrend.csv")

head(Historical_USTrend)
str(Historical_USTrend)

Historical_USTrend$time <- as.Date(Historical_USTrend$time, format = "%Y-%m-%d") 
class(Historical_USTrend$time)
```

```{r US_Total_Trend.plot, fig.cap=paste("Trend of confirmed case in the U.S.")}
US_Total_Trend.plot <-
  ggplot(Historical_USTrend, aes(x = time, y = cum_confirm)) +
  geom_point(colour = "#e6550d") +
  geom_line(colour = "#d95f0e") +
  labs(x = "Time", 
       y = "Confirmed Cases") + 
  scale_x_date(date_labels = "%Y-%m-%d")

print(US_Total_Trend.plot)
```

### Trend of total confirmed case in all states in the U.S.

```{r US_State_Trend.plot, fig.cap=paste("Trend of each states in United States")}
# Create a ggplot depicting cases incereasing over time in each state
US_State_Trend.plot <- ggplot(Covid19_US_processed, aes(x=time, y=cum_confirm, color=state)) +
    geom_line(alpha = 0.95, size = 0.5) +
    geom_text_repel(aes(label=state),
                    function(Covid19_US_processed) 
                      Covid19_US_processed[Covid19_US_processed$time == as.Date("2020-04-16"),])+
    mytheme +
    theme(legend.position = "none") +
    labs(x=expression(paste("Time"))) + 
    labs(y=expression(paste("Comfimed Cases")))+
    labs(color="Covid19_US_processed") +
    scale_x_date(date_labels = "%Y-%m-%d") 

print(US_State_Trend.plot)
```


### Trend of total confirmed case of top ten states in the U.S. (except NY and NJ)
```{r Select top ten states in the U.S. (except NY and NJ), results='hide', message=FALSE, warning=FALSE}
TopTen_States <- Covid19_US_processed %>%
  filter(time >= as.Date("2020-04-16")) %>%
  filter(state!= "New York state" & state!= "New Jersey" & state!= "Unpublishedsources") %>%
  top_n(10, cum_confirm) %>%
  arrange(desc(cum_confirm))

TopTen_States <- pull(TopTen_States, state)

TenStates_US <- Covid19_US_processed %>% 
  filter(state %in% TopTen_States) 

str(TenStates_US)

# Save as csv
write.csv(TenStates_US,
          file = "./Data/Processed/TenStates_US (except NY and NJ).csv")
```


```{r TenStates_Trend.plot, fig.cap=paste("Trend of top ten states in United States (except NY and NJ)")}
# Create a ggplot 
TenStates_Trend.plot <- ggplot(TenStates_US, aes(x = time, y =cum_confirm, color=state)) +
  geom_line(alpha = 0.95, size = 1) +
  geom_text_repel(aes(label=state),
                      function(TenStates_US) 
                      TenStates_US[TenStates_US$time == as.Date("2020-04-16"),])+
  mytheme +
  theme(legend.position = "none") +
  labs(x=expression(paste("Time"))) + 
  labs(y=expression(paste("Comfimed Cases"))) +
  labs(color="state") +
  scale_x_date(date_labels = "%Y-%m-%d")


print(TenStates_Trend.plot)
```

## Analysis trend of dead and heal cases number and rate
### China

```{r Gather Covid19_China_processed dataset, results='hide', message=FALSE, warning=FALSE}
str(Covid19_China_processed)

names(Covid19_China_processed)[1] <- "Province"
names(Covid19_China_processed)[4] <- "Dead Rate"
names(Covid19_China_processed)[6] <- "Heal Rate"

ChinaTrend_gather <- tidyr::gather(Covid19_China_processed, "Type","Rate", 4,6)
head(ChinaTrend_gather)

```


```{r China_Rate.plot, fig.cap=paste("Compare total death rate and heal rate in each province (China)")}
China_Rate.plot <- ggplot(ChinaTrend_gather, aes(fill=Type, y=Rate, x=Province)) + 
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(x=expression(paste("Province Name"))) + 
  labs(y=expression(paste("Rate"))) +
  labs(fill = "Rate Types") +
  mytheme + 
  theme(legend.position = "right") 
print(China_Rate.plot)
```

**Trend of confirms, deaths,and heal number in China**

```{r Gather Historical_China_processed dataset, results='hide', message=FALSE, warning=FALSE}
# Change Historical_China_processed dataset's Column names
str(Historical_China_processed)

Historical_China_processed2 <- Historical_China_processed %>%
  group_by(time,country) %>%
  summarise(total_confirm = sum(cum_confirm),
            total_heal = sum(cum_heal),
            total_dead = sum(cum_dead))
head(Historical_China_processed2)

names(Historical_China_processed2)[1] <- "Time"
names(Historical_China_processed2)[3] <- "Total Confirm"
names(Historical_China_processed2)[4] <- "Total Heal"
names(Historical_China_processed2)[5] <- "Total Dead"

Historical_China_gather <- tidyr::gather(Historical_China_processed2, "Type", "Number", 3:5)
str(Historical_China_gather)
```

```{r China_Number_Trend.plot, fig.cap=paste("Trend of deaths, confirms, and heal number in China")}
China_Number_Trend.plot <- ggplot(Historical_China_gather, aes(Time, Number, color = Type)) +
  geom_point() + 
  geom_line() + 
  labs(x=expression(paste("Time"))) + 
  labs(y=expression(paste("Total number")))+
  mytheme + 
  theme(legend.position = "right") +
  scale_x_date(date_labels = "%Y-%m-%d")  
print(China_Number_Trend.plot)
```

**Trend of confirms, deaths,and heal number in China (12/1/2019-2/15/2020")**

```{r China_Number_Trend_Limit.plot, fig.cap=paste("Trend of deaths, confirms, and heal number in China (12/1/2019-2/15/2019")}
China_Number_Trend_Limit.plot <- ggplot(Historical_China_gather, aes(Time, Number, color = Type)) +
  geom_point() + 
  geom_line() + 
  labs(x=expression(paste("Time"))) + 
  labs(y=expression(paste("Total number")))+
  mytheme + 
  theme(legend.position = "right") +
  scale_x_date(date_labels = "%Y-%m-%d", 
    limits = c(as.Date("2019-12-1"), as.Date("2020-02-15"))) +
  ylim(c(0, 500))  
print(China_Number_Trend_Limit.plot)
```

**Trend of confirms, deaths,and heal number top ten provines, China** 

```{r Gather TenProvinces_China dataset, results='hide', message=FALSE, warning=FALSE}
head(TenProvinces_China)

names(TenProvinces_China)[1] <- "Time"
names(TenProvinces_China)[3] <- "Total Confirm"
names(TenProvinces_China)[4] <- "Total Heal"
names(TenProvinces_China)[5] <- "Total Dead"

TenProvinces_China_gather <- tidyr::gather(TenProvinces_China, "Type", "Number", 3:5)
str(TenProvinces_China_gather)
```


```{r TenProvinces_Number_Trend.plot (expect Hubei), fig.cap=paste("Trend of deaths, confirms, and heal number in top ten provine (expect Hubei), China")}
TenProvinces_Number_Trend.plot <-
  ggplot(TenProvinces_China_gather, aes(x = Time, y = Number, color = Type)) +
  geom_line() +
  geom_point() +
  labs(x=expression(paste("Time"))) + 
  labs(y=expression(paste("Total number")))+
  mytheme + 
  theme(legend.position = "right") +
  scale_x_date(date_labels = "%Y-%m-%d", 
    limits = c(as.Date("2020-1-15"), as.Date("2020-02-15"))) +
  ylim(c(0, 500)) +
  facet_wrap(.~province)
print(TenProvinces_Number_Trend.plot)

```

### Hubei province, China

**Trend of confirms, deaths,and heal number in top five cities, Hubei(12/1/2019-3/15/2020)**

```{r Select and gather top five cities (Hubei), results='hide', message=FALSE, warning=FALSE}
head(Historical_Hubei)

# Select top five cities
TopFive_Hubei <- Historical_Hubei %>%
  filter(time >= as.Date("2020-04-16")) %>%
  top_n(5, cum_confirm) %>%
  arrange(desc(cum_confirm))

TopFive_Hubei <- pull(TopFive_Hubei, city)

TopFiveCity_Hubei <- filter(Historical_Hubei, city %in% TopFive_Hubei) %>%
  arrange(desc(cum_confirm))

head(TopFiveCity_Hubei)

# Change column names
names(TopFiveCity_Hubei)[1] <- "Time"
names(TopFiveCity_Hubei)[5] <- "Total Confirm"
names(TopFiveCity_Hubei)[6] <- "Total Heal"
names(TopFiveCity_Hubei)[7] <- "Total Dead"

Historical_Hubei_gather <- tidyr::gather(TopFiveCity_Hubei, "Type", "Number", 5:7)
head(Historical_Hubei_gather)
```


```{r FiveCities_Number_Trend.plot, fig.cap=paste("Trend of confirms, deaths,and heal number in top five cities, Hubei(12/1/2019-3/15/2020)")}
FiveCities_Number_Trend.plot <-
  ggplot(Historical_Hubei_gather, aes(x = Time, y = Number, color = Type)) +
  geom_line() +
  geom_point() +
  labs(x=expression(paste("Time"))) + 
  labs(y=expression(paste("Total number")))+
  mytheme + 
  theme(legend.position = "right") +
  scale_x_date(date_labels = "%Y-%m-%d", 
    limits = c(as.Date("2019-12-1"), as.Date("2020-03-15"))) +
  ylim(c(0, 5000)) +
  facet_wrap(.~city)
print(FiveCities_Number_Trend.plot)

```



### United States

```{r Gather Historical_USTrend dataset, results='hide', message=FALSE, warning=FALSE}
# Change Historical_USTrend dataset's Column names
str(Historical_USTrend)

Historical_USTrend2 <- Historical_USTrend
names(Historical_USTrend2)[1] <- "Time"
names(Historical_USTrend2)[3] <- "Total Confirm"
names(Historical_USTrend2)[4] <- "Total Heal"
names(Historical_USTrend2)[5] <- "Total Dead"

USTrend_gather <- tidyr::gather(Historical_USTrend2, "Type", "Number", 3:5)
str(USTrend_gather)
```


```{r US_Number_Trend.plot, fig.cap=paste("Trend of confirms, deaths,and heal number in the U.S.")}
US_Number_Trend.plot <- ggplot(USTrend_gather, aes(Time, Number, color = Type)) +
  geom_point() + 
  geom_line() + 
  labs(x=expression(paste("Time"))) + 
  labs(y=expression(paste("Total number")))+
  mytheme + 
  theme(legend.position = "right") +
  scale_x_date(date_labels = "%Y-%m-%d", 
    limits = c(as.Date("2020-01-15"), as.Date("2020-04-16"))) 
print(US_Number_Trend.plot)


```


**Heal and dead plot for top ten states**

```{r Gather TenStates_US dataset, results='hide', message=FALSE, warning=FALSE}
head(TenStates_US)

names(TenStates_US)[1] <- "Time"
names(TenStates_US)[4] <- "Total Confirm"
names(TenStates_US)[5] <- "Total Heal"
names(TenStates_US)[6] <- "Total Dead"

TenStates_US_gather <- tidyr::gather(TenStates_US, "Type", "Number", 4:6)
str(TenStates_US_gather)
```


```{r Ten_States_Number_Trend.plot, fig.cap=paste("Trend of confirms, deaths,and heal number in top ten states")}
Ten_States_Number_Trend.plot <-
  ggplot(TenStates_US_gather, aes(x = Time, y = Number, color = Type)) +
  geom_line() +
  geom_point() +
  labs(x=expression(paste("Time"))) + 
  labs(y=expression(paste("Total number")))+
  mytheme + 
  theme(legend.position = "right") +
  facet_wrap(.~state)
print(Ten_States_Number_Trend.plot)

```
