---
title: "Data exploration"
author: "Xueying Feng"
date: "4/11/2020"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(lubridate)
require(dplyr)

```

## Set theme

```{r}
require(ggplot2)
#install.packages("ggrepel") 
require(ggrepel)

mytheme <- theme_minimal(base_size = 12, base_family = "Times") + 
  theme(axis.text.x = element_text(color = "DarkGrey"),
        legend.position = "top") 
theme_set(mytheme)

```

## Increasing case trend in China

```{r, fig.cap=paste("Trend of confirmed case in China from Dec-2019 to Apri-2020")}
#Import datasets
Historical_Global_processed <- read.csv("./Data/Processed/Historical_Global_processed.csv")

#Selsct data from China
Historical_China = Historical_Global_processed[Historical_Global_processed$country == 'China',]
head(Historical_China)

# Change date column to date
Historical_China$time <- as.Date(Historical_China$time, format = "%Y-%m-%d") 
class(Historical_China$time)

# Create a ggplot depicting cases incereasing over time
China_Trend <- 
  ggplot(Historical_China, aes(x=time, y=cum_confirm)) +
  geom_point(colour = "#e6550d") +
  geom_line(colour = "#d95f0e") +
  labs(x = "Time", 
       y = "Confirmed Cases") + 
  scale_x_date(date_labels = "%Y-%m-%d")

print(China_Trend)

```

## Compare total number in each province 

```{r, fig.cap=paste("Compare total number in each province, China")}
Covid19_China_processed <- read.csv("./Data/Processed/Covid19_China_processed.csv")

# scale_y_log10: transform the y-axis to make it easier to read
China.plot <-ggplot(Covid19_China_processed,aes(x = name,y = confirm)) + 
  geom_bar(size = 2, stat = "identity",position = "dodge", fill ="#fdae6b") + 
  scale_y_log10() +
  labs(x= "Province Name (China)",
       y = "Numbers of Confirmed Cases") + 
  mytheme +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

print(China.plot)
```

## Analysis of trend of each province in China 

```{r}
#Import datasets
Historical_China_processed <- read.csv("./Data/Processed/Historical_China_processed.csv",header = TRUE)

head(Historical_China_processed)
str(Historical_China_processed)

# Change date column to date
Historical_China_processed$time <- as.Date(Historical_China_processed$time, format = "%m/%d/%y") 
head(Historical_China_processed)
class(Historical_China_processed$time)

```

## Only select Hubei data

```{r}
# Select Hubei province data
Historical_Hubei = Historical_China_processed[Historical_China_processed$province == 'Hubei',]
head(Historical_Hubei)

# Group by Hubei cities
Historical_HubeiCityV2 <- Historical_Hubei %>%
  filter(city!= "Location TBD") %>%
  group_by(time,city) %>%
  summarise(total_confirm = sum(cum_confirm))
```

## Draw trend line of each city in Hubei province
```{r, fig.cap=paste("Trend of each city in Hubei province, China")}
# Create a ggplot depicting cases incereasing over time
HubeiCity_Trend <- ggplot(Historical_HubeiCityV2, aes(x=time, y=total_confirm, color=city)) +
    geom_line(alpha = 0.95, size = 0.5) +
    geom_text_repel(aes(label=city),
                    function(Historical_HubeiCityV2) 
                      Historical_HubeiCityV2[Historical_HubeiCityV2$time == as.Date("2020-04-16"),])+
    mytheme +
    theme(legend.position = "none") +
    labs(x=expression(paste("Time"))) + 
    labs(y=expression(paste("Comfimed Cases")))+
    labs(color="city") +
    scale_x_date(date_labels = "%Y-%m-%d")

print(HubeiCity_Trend)
```

## Draw trend line of each city in Hubei province (except Wuhan)
```{r, fig.cap=paste("Trend of each city in Hubei province (except Wuhan)")}
# Group by Hubei cities
Historical_HubeiCityV3 <- Historical_Hubei %>%
  filter(city!= "Location TBD" & city!= "Wuhan") %>%
  group_by(time,city) %>%
  summarise(total_confirm = sum(cum_confirm))

# Create a ggplot depicting cases incereasing over time
HubeiCity_Trend <- ggplot(Historical_HubeiCityV3, aes(x=time, y=total_confirm, color=city)) +
    geom_line(alpha = 0.95, size = 0.5) +
    geom_text_repel(aes(label=city),
                    function(Historical_HubeiCityV3) 
                      Historical_HubeiCityV3[Historical_HubeiCityV3$time == as.Date("2020-04-16"),])+
    mytheme +
    theme(legend.position = "none") +
    labs(x=expression(paste("Time"))) + 
    labs(y=expression(paste("Comfimed Cases")))+
    labs(color="city") +
    scale_x_date(date_labels = "%Y-%m-%d")

print(HubeiCity_Trend)
```

## Remove HUbei Data

```{r}
# Remove Hubei province data
Historical_ChinaProvince = Historical_China_processed[Historical_China_processed$province != 'Hubei',]
head(Historical_ChinaProvince)

Historical_ChinaProvinceV2 <- Historical_ChinaProvince %>%
  group_by(time,province) %>%
  summarise(total_confirm = sum(cum_confirm))
```

## Draw trend line of each province (except Hubei)
```{r, fig.cap=paste("Trend of each province, China (except Hubei)")}
# Create a ggplot depicting cases incereasing over time
ChinaProvince_Trend <- ggplot(Historical_ChinaProvinceV2, aes(x=time, y=total_confirm, color=province)) +
    geom_line(alpha = 0.95, size = 1) +
    geom_text_repel(aes(label=province),
                    function(Historical_ChinaProvinceV2) 
                      Historical_ChinaProvinceV2[Historical_ChinaProvinceV2$time == as.Date("2020-04-16"),])+
    mytheme +
    theme(legend.position = "none") +
    labs(x=expression(paste("Time"))) + 
    labs(y=expression(paste("Comfimed Cases"))) +
    labs(color="province") +
    scale_x_date(date_labels = "%Y-%m-%d")

print(ChinaProvince_Trend)
```

## Top ten province in China (except Hubei)
```{r}
# Filter out the top ten provinces with the highest number of diagnoses (except Hubei)
TopTen_provinces <- Historical_ChinaProvinceV2 %>% 
  filter(time >= as.Date("2020-04-16")) %>%
  top_n(10, total_confirm) %>%
  arrange(desc(total_confirm)) 

TopTen_provinces <- pull(TopTen_provinces, province)

TenProvinces_China <- filter(Historical_ChinaProvinceV2, province %in% TopTen_provinces) %>%
  arrange(desc(total_confirm))
TopTen_provinces
```

## Draw trend line of top ten province (except Hubei)
```{r, fig.cap=paste("Trend of top ten provinces, China (except Hubei)")}
# Draw plot
TenProvince_trend <- ggplot(TenProvinces_China, aes(x = time, y =total_confirm, color=province)) +
    geom_line(alpha = 0.95, size = 1) +
    geom_text_repel(aes(label=province),
                    function(Historical_ChinaProvinceV2) 
                      Historical_ChinaProvinceV2[Historical_ChinaProvinceV2$time == as.Date("2020-04-16"),])+
    mytheme +
    theme(legend.position = "none") +
    labs(x=expression(paste("Time"))) + 
    labs(y=expression(paste("Comfimed Cases"))) +
    labs(color="province") +
    scale_x_date(date_labels = "%Y-%m-%d")

print(TenProvince_trend)
```


## Compare total number in each country 

```{r, fig.cap=paste("Compare total number in each country")}
Covid19_Global_processed <- read.csv("./Data/Processed/Covid19_Global_processed.csv")

# Filter out top 15 countries with the highest number of diagnoses
Fifty_countries <- Covid19_Global_processed %>% 
  top_n(15, confirm) %>%
  arrange(desc(confirm))

# Draw histogram plot
Global.plot <-ggplot(Fifty_countries,aes(x = name,y = confirm)) + 
  geom_bar(size = 2, stat = "identity",position = "dodge", fill ="#fdae6b") + 
  labs(x = " Country Name (Top 15)",
       y = "Numbers of Confirmed Cases") + 
  mytheme +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

print(Global.plot)
```

## TopTen_countries (global level except China)

```{r cars}
# Check on data frame
head(Historical_Global_processed)
str(Historical_Global_processed)

# Change date column to date
Historical_Global_processed$time <- as.Date(Historical_Global_processed$time, format = "%Y-%m-%d") 
class(Historical_Global_processed$time)

# Remove China data
Historical_Global = Historical_Global_processed[Historical_Global_processed$country != 'China',]
head(Historical_Global)

# Filter out the top ten countries with the highest number of diagnoses (except China)
TopTen_countries <- Historical_Global %>% 
  filter(time >= as.Date("2020-04-16")) %>%
  top_n(10, cum_confirm) %>%
  arrange(desc(cum_confirm)) 

# Selects a column in a data frame and transforms it into a vector
TopTen_countries <- pull(TopTen_countries, country)

Historical_TopTen <- filter(Historical_Global, country %in% TopTen_countries) %>%
  arrange(desc(cum_confirm))
str(Historical_TopTen)

```

## Plots trends of comfirmed cases in top ten coutries (except China)

```{r, fig.cap=paste("Trend of Top Ten Countries (except China)")}

# Draw plot
TenCountries <- ggplot(Historical_TopTen, aes(x = time, y =cum_confirm, color=country)) +
    geom_line(alpha = 0.95, size = 1) +
    geom_text_repel(aes(label=country),
                    function(Historical_TopTen) 
                      Historical_TopTen[Historical_TopTen$time == as.Date("2020-04-16"),])+
    mytheme +
    theme(legend.position = "none") +
    labs(x=expression(paste("Time"))) + 
    labs(y=expression(paste("Comfimed Cases"))) +
    labs(color="Country") +
    scale_x_date(date_labels = "%Y-%m-%d")

print(TenCountries)
```


