---
output: 
  pdf_document:
    keep_tex: yes
    fig_caption: yes
    number_sections: yes
geometry: margin=2.54cm
title: "Covid-19 Trend Analysis and Visualization"
subtitle: "https://github.com/xueying-F/ENV872-Project_XY_Feng"
author: "Xueying Feng"
fontsize: 12pt
mainfont: Times New Roman
editor_options: 
  chunk_output_type: console
---

\newpage
\tableofcontents 
\newpage
\listoftables 
\newpage
\listoffigures 
\newpage

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Set your working directory
getwd()
setwd("~/Desktop/Environ 872/ENV872 Project_XY_Feng")

# Load your packages
library(plyr) #Tools for Splitting, Applying and Combining Data
library(tidyverse) #Getting data frames to tidy
library(lubridate) #For dates and date-times

require(ggplot2)
#install.packages("ggrepel") 
require(ggrepel)

#install.packages("usmap")
library(usmap)

# Set your ggplot theme
mytheme <- theme_minimal(base_size = 12, base_family = "Times") + 
  theme(axis.text.x = element_text(color = "DarkGrey"),
        legend.position = "top") 
theme_set(mytheme)

# Load your datasets
Historical_Global_processed <- read.csv("./Data/Processed/Historical_Global_processed.csv")
Historical_China_processed <- read.csv("./Data/Processed/Historical_China_processed.csv")
Covid19_Global_processed <- read.csv("./Data/Processed/Covid19_Global_processed.csv")
Covid19_China_processed <- read.csv("./Data/Processed/Covid19_China_processed.csv")
Covid19_US_processed <- read.csv("./Data/Processed/Covid19_US_processed.csv")

```

# Rationale and Research Questions

China broke out a global disease, COVID-19, in December 2019. It has been confirmed that it is a new type of virus called 2019 new coronavirus (2019-nCoV) in Jan 2020. The World Health Organization (WHO) announced the official new name of the disease caused by nCoV2019 (2019 novel coronavirus) in Feb 2020. The CDC noted that the symptoms of the new coronavirus include "symptoms of fever and lower respiratory tract disease (eg. coughing, difficulty breathing). There is currently no vaccine, and because this is a virus, antibiotics will not work. 

COVID-19 is a global pandemic disease confirmed by the World Health Organization (WHO), and likely to occur when a new virus appears. Also, it can easily infect humans, and can spread from person to person in a rapid and sustained manner. This research focus on the datasets about China, United States, and whole global level to see the increase trend of confirmed cases, death cases and heal cases. Moreover, I will combine some news report that I read to analyze the trend pattern.

The research question is:
* Is there a same trend pattern on China and United States? 
* What is trend pattern on global level?
* Why trend patterns look like that?

Downloading datasets from R package, nCov2019 (https://github.com/GuangchuangYu/nCov2019), which provides convenient access to epidemiological data on the coronavirus outbreak, which contains real-time data and historical data for each country. Please see the detail information in Dataset Information setion.


\newpage

# Dataset Information

## Database Information
I access all Novel Coronavirus data from the R package, nCov2019, which includes detailed real-time statistics, historical data in all countries, and down to the city-level.

More information can be found here:https://github.com/GuangchuangYu/nCov2019.


## Data Content Information

I pulled all relevent data from nCov2019 package. I named all pulled data as raw.csv and then saved into Data/Raw folder. Beside, I wrangled all raw and selected the relevent columns, and then saved into Data/Processed folder. Because data pulled from nCov2019 package are real-time data, I kept the update time on April 16th, 2020. Therefore, the following descriptions of my datasets are processed data.

```{r, table1, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
#install.packages("pander")
require(pander)
panderOptions('table.split.table', Inf)
set.caption("Historical_Global_processed dataset")
my.data <- "  Column name | Description 
 time | Date 
 country | Country name 
 cum_confirm | Cumulative number of COVID-19 confirmed cases  
 cum_heal | Cumulative number of COVID-19 heal cases 
 cum_dead | Cumulative number of COVID-19 death cases "
df <- read.delim(textConnection(my.data),header=FALSE,sep="|",strip.white=TRUE,stringsAsFactors=FALSE)
names(df) <- unname(as.list(df[1,])) # put headers on
df <- df[-1,] # remove first row
row.names(df)<-NULL
pander(df, style = 'rmarkdown')
```

```{r, table2, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
panderOptions('table.split.table', Inf)
set.caption("Historical_China_processed dataset")
my.data <- "  Column name | Description 
 time | Date 
 country | Country name 
 province | Province name 
 city | City name 
 cum_confirm | Cumulative number of COVID-19 confirmed cases  
 cum_heal | Cumulative number of COVID-19 heal cases 
 cum_dead | Cumulative number of COVID-19 death cases "
df <- read.delim(textConnection(my.data),header=FALSE,sep="|",strip.white=TRUE,stringsAsFactors=FALSE)
names(df) <- unname(as.list(df[1,])) # put headers on
df <- df[-1,] # remove first row
row.names(df)<-NULL
pander(df, style = 'rmarkdown')
```

```{r, table3, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
panderOptions('table.split.table', Inf)
set.caption("Covid19_Global_processed dataset")
my.data <- "  Column name | Description 
 name | Country name 
 confirm | Number of COVID-19 confirmed cases 
 dead | Number of COVID-19 death cases 
 deadRate | Total death number/ Cumulative total number of COVID-19 cases(%)
 heal | number of COVID-19 heal cases 
 healRate | Total heal number/ Cumulative total number of COVID-19 cases(%)"
df <- read.delim(textConnection(my.data),header=FALSE,sep="|",strip.white=TRUE,stringsAsFactors=FALSE)
names(df) <- unname(as.list(df[1,])) # put headers on
df <- df[-1,] # remove first row
row.names(df)<-NULL
pander(df, style = 'rmarkdown')
```

```{r, table4, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
panderOptions('table.split.table', Inf)
set.caption("Covid19_China_processed dataset")
my.data <- "  Column name | Description 
 name | Province name 
 confirm | Number of COVID-19 confirmed cases 
 dead | Number of COVID-19 death cases 
 deadRate | Total death number/ Cumulative total number of COVID-19 cases(%)
 heal | number of COVID-19 heal cases 
 healRate | Total heal number/ Cumulative total number of COVID-19 cases(%) "
df <- read.delim(textConnection(my.data),header=FALSE,sep="|",strip.white=TRUE,stringsAsFactors=FALSE)
names(df) <- unname(as.list(df[1,])) # put headers on
df <- df[-1,] # remove first row
row.names(df)<-NULL
pander(df, style = 'rmarkdown')
```

```{r, table5, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
panderOptions('table.split.table', Inf)
set.caption("Covid19_US_processed dataset")
my.data <- "  Column name | Description 
 time | Date |
 country | United States 
 province | States name 
 cum_confirm | Cumulative number of COVID-19 confirmed cases  
 cum_heal | Cumulative number of COVID-19 heal cases 
 cum_dead | Cumulative number of COVID-19 death cases "
df <- read.delim(textConnection(my.data),header=FALSE,sep="|",strip.white=TRUE,stringsAsFactors=FALSE)
names(df) <- unname(as.list(df[1,])) # put headers on
df <- df[-1,] # remove first row
row.names(df)<-NULL
pander(df, style = 'rmarkdown')
```

\newpage


# Exploratory Analysis 

Wrangling the raw data for thess datasets is to select the columns that are useful for this research. After wrangling all dataset, they are saved into processed folder. However, this section will show how processed data form, but processed data will be directly used to do analysis. During analysis, datasets will be wrangled again based on anlysis requirments.

```{r}
#Covid19_China_raw <- read.csv("Covid19_China_raw.csv",
#                              header=TRUE,stringsAsFactors = FALSE, strip.white = TRUE,sep = ',')
#Covid19_China_processed <- 
#  select(Covid19_China_raw, name, confirm, dead, deadRate, heal, healRate)


#Covid19_US_raw <- read.csv("Covid19_US_raw.csv",header=TRUE,stringsAsFactors = FALSE, strip.white = TRUE,sep = ',')
#Covid19_US_processed <- 
#  select(Covid19_US_raw, time:cum_dead)


#Covid19_Global_raw <- read.csv("Covid19_Global_raw.csv",
#                               header=TRUE,stringsAsFactors = FALSE, strip.white = TRUE,sep = ',')
#Covid19_Global_processed <- 
#  select(Covid19_Global_raw, name, confirm, dead, deadRate,heal, healRate)


#historical_China_raw <- read.csv("historical_China_raw.csv",
#                                 header=TRUE,stringsAsFactors = FALSE, strip.white = TRUE,sep = ',')
#Historical_China_processed <- 
#  select(historical_China_raw, time:cum_dead)


#historical_Global_raw <- read.csv("historical_Global_raw.csv",header=TRUE,stringsAsFactors = FALSE, strip.white = TRUE,sep = ',')
#Historical_Global_processed <- historical_Global_raw
```


## Explore China Dataset

```{r, results='hide', message=FALSE, warning=FALSE}
Covid19_China_processed <- read.csv("./Data/Processed/Covid19_China_processed.csv")

# Check data frame
colnames(Covid19_China_processed)
head(Covid19_China_processed)
```

```{r, echo =FALSE, fig.cap="Compare total confimed number in each province in China"}
# scale_y_log10: transform the y-axis to make it easier to read
China.plot <-ggplot(Covid19_China_processed,aes(x = name,y = confirm)) + 
  geom_bar(size = 2, stat = "identity",position = "dodge", fill ="#fdae6b") + 
  scale_y_log10() +
  labs(x= "Province Name (China)",
       y = "Numbers of Confirmed Cases") + 
  mytheme +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

print(China.plot)
```


## Explore Global Dataset 

**Top 15 countries**
```{r, results='hide', message=FALSE, warning=FALSE}
Covid19_Global_processed <- read.csv("./Data/Processed/Covid19_Global_processed.csv")

# Check data frame
head(Historical_Global_processed)
str(Historical_Global_processed)

# Change date column to date
Historical_Global_processed$time <- as.Date(Historical_Global_processed$time, format = "%Y-%m-%d") 
class(Historical_Global_processed$time)


# Filter out top 15 countries with the highest number of diagnoses
Fifty_countries <- Covid19_Global_processed %>% 
  top_n(15, confirm) %>%
  arrange(desc(confirm))

# Save as csv
write.csv(Fifty_countries,
          file = "./Data/Processed/Top_Fifty_countries (cum).csv", row.names=FALSE)
```


```{r, fig.cap=paste("Compare total confimed number of each country"),echo = FALSE}
# Draw histogram plot
Global.plot <-ggplot(Fifty_countries,aes(x = name,y = confirm)) + 
  geom_bar(size = 2, stat = "identity",position = "dodge", fill ="#fdae6b") + 
  labs(x = " Country Name (Top 15)",
       y = "Numbers of Confirmed Cases") + 
  mytheme +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

print(Global.plot)
```

Figure 2 shows the total confimed number of each country (update on April 16th, 2020). Due to this graph, I will focus on the United States dataset and see its trend pattern.


## Explore United States Dataset

**Map of comfimed cases didtribution in United State**

```{r, results='hide', message=FALSE, warning=FALSE}
# Check data
head(Covid19_US_processed)
str(Covid19_US_processed)

# Change date column to date
Covid19_US_processed$time <- as.Date(Covid19_US_processed$time, format = "%m/%d/%y") 
class(Covid19_US_processed$time)

```

```{r, results='hide', message=FALSE, warning=FALSE}
# Check column names
colnames(Covid19_US_processed)

# Rename column where names is "province"
names(Covid19_US_processed)[names(Covid19_US_processed) == "province"] <- "state"
str(Covid19_US_processed)


# Select total confirmed cases in 51 States
US_States <- Covid19_US_processed %>%
  filter(time == as.Date("2020-04-16")) %>%
  filter(state!= "Unpublished sources") %>%
  top_n(51, cum_confirm)

# Convert state column to characters 
US_States$state <- as.character(US_States$state)

# Rename specific cell name
US_States$state[US_States$state == 'New York state'] <- 'New York'
US_States$state[US_States$state == "Washington State"] <- "Washington"
US_States$state[US_States$state == 'the state of Wisconsin'] <- 'Wisconsin'


```

```{r, fig.cap=paste("Map of comfimed cases didtribution in United State"), echo = FALSE}
#install.packages("usmap")
library(usmap)

USMap_Base <- plot_usmap(regions = "counties") + 
  labs(title = "US Counties",
       subtitle = "This is a blank map of the counties of the United States.") + 
  theme(panel.background = element_rect(color = "black", fill = "lightblue"))


### Draw US Case distribution on map
CasesDidtribution <- plot_usmap(regions="state", data = US_States, values = "cum_confirm", color = "black") +
  scale_fill_continuous(low = "#fee6ce", high = "#d95f0e", 
                        name = "Total confirmed cases", label = scales::comma) + 
  labs(title = "State Reporting Cases of Covid 19") +
  theme(legend.position = "right")

print(CasesDidtribution)
```

* Figure 1 shows the total confimed number of each province in China (update on April 16th, 2020). Due to this graph, I will seperate Hubei province and other provinces in China and see their trend pattern.

* Figure 2 shows the total confimed number of each country (update on April 16th, 2020). Due to this graph, I will focus on the United States dataset and see its trend pattern.

* Figure 3 shows confirmed cases distribution in each state by using map, which can make readers straighter forward to see which states are more serious. 

\newpage

# Analysis

## Anlysis trend of confirmed cases in China
### Trend of Total confirmed cases in China
```{r, fig.cap=paste("Trend of confirmed case in China"), echo = FALSE}
# Selsct data from China
str(Historical_Global_processed)
Historical_China = Historical_Global_processed[Historical_Global_processed$country == 'China',]
head(Historical_China)

# Create a ggplot depicting cases incereasing over time
China_Trend <- 
  ggplot(Historical_China, aes(x=time, y=cum_confirm)) +
  geom_point(colour = "#e6550d") +
  geom_line(colour = "#d95f0e") +
  labs(x = "Time", 
       y = "Confirmed Cases") + 
  scale_x_date(date_labels = "%Y-%m-%d")

print(China_Trend)
```


### Trend of confirmed cases in each province in China 

```{r, results='hide', message=FALSE, warning=FALSE}
# Check data
head(Historical_China_processed)
str(Historical_China_processed)

# Change date column to date
Historical_China_processed$time <- as.Date(Historical_China_processed$time, format = "%m/%d/%y") 
head(Historical_China_processed)
class(Historical_China_processed$time)

```

### Trend of confirmed cases in Hubei, China

```{r, results='hide', message=FALSE, warning=FALSE}
# Select Hubei province data
Historical_Hubei = Historical_China_processed[Historical_China_processed$province == 'Hubei',]
head(Historical_Hubei)

# Group by Hubei cities
Historical_HubeiCityV2 <- Historical_Hubei %>%
  select(time, city:cum_dead) 

# Save as csv
write.csv(Historical_HubeiCityV2,
          file = "./Data/Processed/Historical_HubeiCity.csv",row.names=FALSE)
```

```{r, fig.cap=paste("Trend of each city in Hubei province, China"), echo = FALSE}
# Create a ggplot depicting cases incereasing over time
HubeiCity_Trend <- ggplot(Historical_HubeiCityV2, aes(x=time, y=cum_confirm, color=city)) +
    geom_line(alpha = 0.95, size = 0.5) +
    geom_text_repel(aes(label=city),
                    function(Historical_HubeiCityV2) 
                      Historical_HubeiCityV2[Historical_HubeiCityV2$time == as.Date("2020-04-16"),])+
    mytheme +
    theme(legend.position = "none") +
    labs(x=expression(paste("Time"))) + 
    labs(y=expression(paste("Comfimed Cases")))+
    labs(color="city") +
    scale_x_date(date_labels = "%Y-%m-%d")

print(HubeiCity_Trend)
```


### Trend of confirmed cases in Hubei cities, China (except Wuhan)

```{r, results='hide', message=FALSE, warning=FALSE}
# Group by Hubei cities
Historical_HubeiCityV3 <- Historical_HubeiCityV2 %>%
  filter(city!= "Location TBD" & city!= "Wuhan") 

# Save as csv
write.csv(Historical_HubeiCityV3,
          file = "./Data/Processed/Historical_HubeiCity (except Wuhan).csv", row.names=FALSE)
```

```{r, fig.cap=paste("Trend of each city in Hubei province,China (except Wuhan)"),echo = FALSE}
# Create a ggplot depicting cases incereasing over time
HubeiCity_Trend <- ggplot(Historical_HubeiCityV3, aes(x=time, y=cum_confirm, color=city)) +
    geom_line(alpha = 0.95, size = 0.5) +
    geom_text_repel(aes(label=city),
                    function(Historical_HubeiCityV3) 
                      Historical_HubeiCityV3[Historical_HubeiCityV3$time == as.Date("2020-04-16"),])+
    mytheme +
    theme(legend.position = "none") +
    labs(x=expression(paste("Time"))) + 
    labs(y=expression(paste("Comfimed Cases")))+
    labs(color="city") +
    scale_x_date(date_labels = "%Y-%m-%d")

print(HubeiCity_Trend)
```


### Trend of confirmed cases in China provinces(except Hubei)

```{r, results='hide', message=FALSE, warning=FALSE}
# Remove Hubei province data
Historical_ChinaProvince = Historical_China_processed[Historical_China_processed$province != 'Hubei',]
head(Historical_ChinaProvince)

Historical_ChinaProvinceV2 <- Historical_ChinaProvince %>%
  group_by(time,province) %>%
  summarise(total_confirm = sum(cum_confirm),
            total_heal = sum(cum_heal),
            total_dead = sum(cum_dead)) 

# Save as csv
write.csv(Historical_ChinaProvinceV2,
          file = "./Data/Processed/Historical_ChinaProvince (except Hubei).csv", row.names=FALSE)
```

```{r, fig.cap=paste("Trend of each province in China (except Hubei)"),echo = FALSE}
# Create a ggplot depicting cases incereasing over time
ChinaProvince_Trend <- ggplot(Historical_ChinaProvinceV2, aes(x=time, y=total_confirm, color=province)) +
    geom_line(alpha = 0.95, size = 1) +
    geom_text_repel(aes(label=province),
                    function(Historical_ChinaProvinceV2) 
                      Historical_ChinaProvinceV2[Historical_ChinaProvinceV2$time == as.Date("2020-04-16"),])+
    mytheme +
    theme(legend.position = "none") +
    labs(x=expression(paste("Time"))) + 
    labs(y=expression(paste("Comfimed Cases"))) +
    labs(color="province") +
    scale_x_date(date_labels = "%Y-%m-%d")

print(ChinaProvince_Trend)
```


### Trend of confirmed cases of top ten provinces in China (except Hubei)
```{r, results='hide', message=FALSE, warning=FALSE}
# Filter out the top ten provinces with the highest number of diagnoses (except Hubei)
TopTen_provinces <- Historical_ChinaProvinceV2 %>% 
  filter(time >= as.Date("2020-04-16")) %>%
  top_n(10, total_confirm) %>%
  arrange(desc(total_confirm)) 

TopTen_provinces <- pull(TopTen_provinces, province)

TenProvinces_China <- filter(Historical_ChinaProvinceV2, province %in% TopTen_provinces) %>%
  arrange(desc(total_confirm))

head(TenProvinces_China)

# Save as csv
write.csv(TenProvinces_China,
          file = "./Data/Processed/TenProvinces_China (except Hubei).csv", row.names=FALSE)
```

```{r, fig.cap=paste("Trend of top ten provinces, China (except Hubei)"), echo = FALSE}
# Draw plot
TenProvince_trend <- ggplot(TenProvinces_China, aes(x = time, y =total_confirm, color=province)) +
    geom_line(alpha = 0.95, size = 1) +
    geom_text_repel(aes(label=province),
                    function(Historical_ChinaProvinceV2) 
                      Historical_ChinaProvinceV2[Historical_ChinaProvinceV2$time == as.Date("2020-04-16"),])+
    mytheme +
    theme(legend.position = "none") +
    labs(x=expression(paste("Time"))) + 
    labs(y=expression(paste("Comfimed Cases"))) +
    labs(color="province") +
    scale_x_date(date_labels = "%Y-%m-%d")

print(TenProvince_trend)
```


## Anlysis Global trend of confirmed cases of top ten countries (except China) 

```{r cars}
# Remove China data
Historical_Global = Historical_Global_processed[Historical_Global_processed$country != 'China',]
str(Historical_Global)

# Filter out the top ten countries with the highest number of diagnoses (except China)
TopTen_countries <- Historical_Global %>% 
  filter(time >= as.Date("2020-04-16")) %>%
  top_n(10, cum_confirm) %>%
  arrange(desc(cum_confirm)) 

# Selects a column in a data frame and transforms it into a vector
TopTen_countries <- pull(TopTen_countries, country)

Historical_TopTen <- filter(Historical_Global, country %in% TopTen_countries) %>%
  arrange(desc(cum_confirm))
head(Historical_TopTen)

# Save as csv
write.csv(Historical_TopTen,
          file = "./Data/Processed/TopTen_countries_trend(except China).csv", row.names=FALSE)
```

```{r, fig.cap=paste("Trend of Top Ten Countries (except China)"), echo = FALSE}
# Draw plot
TenCountries <- ggplot(Historical_TopTen, aes(x = time, y =cum_confirm, color=country)) +
    geom_line(alpha = 0.95, size = 1) +
    geom_text_repel(aes(label=country),
                    function(Historical_TopTen) 
                      Historical_TopTen[Historical_TopTen$time == as.Date("2020-04-16"),])+
    mytheme +
    theme(legend.position = "none") +
    labs(x=expression(paste("Time"))) + 
    labs(y=expression(paste("Comfimed Cases"))) +
    labs(color="Country") +
    scale_x_date(date_labels = "%Y-%m-%d")

print(TenCountries)
```


## Analysis trend of confirmed cases in United States. 

```{r,results='hide', message=FALSE, warning=FALSE}
Historical_USTrend <- Historical_Global_processed[Historical_Global_processed$country == 'United States',] %>%
  na.omit(Historical_USTrend)

tail(Historical_USTrend)

# Save as csv
write.csv(Historical_USTrend,
          file = "./Data/Processed/Historical_USTrend.csv",row.names=FALSE)

```

### Trend of total confirmed case in United States.

```{r, results='hide', message=FALSE, warning=FALSE}
Historical_USTrend <- read.csv("./Data/Processed/Historical_USTrend.csv")

head(Historical_USTrend)
str(Historical_USTrend)

Historical_USTrend$time <- as.Date(Historical_USTrend$time, format = "%Y-%m-%d") 
class(Historical_USTrend$time)
```

```{r, fig.cap=paste("Trend of confirmed case in the U.S."), echo=FALSE}
US_ComfirmedCases_Trend <-
  ggplot(Historical_USTrend, aes(x = time, y = cum_confirm)) +
  geom_point(colour = "#e6550d") +
  geom_line(colour = "#d95f0e") +
  labs(x = "Time", 
       y = "Confirmed Cases") + 
  scale_x_date(date_labels = "%Y-%m-%d")

print(US_ComfirmedCases_Trend)
```

### Trend of total confirmed case in all states in the U.S.

```{r, fig.cap=paste("Trend of each states in United States"), echo = FALSE}
# Create a ggplot depicting cases incereasing over time in each state
US_State_Trend <- ggplot(Covid19_US_processed, aes(x=time, y=cum_confirm, color=state)) +
    geom_line(alpha = 0.95, size = 0.5) +
    geom_text_repel(aes(label=state),
                    function(Covid19_US_processed) 
                      Covid19_US_processed[Covid19_US_processed$time == as.Date("2020-04-16"),])+
    mytheme +
    theme(legend.position = "none") +
    labs(x=expression(paste("Time"))) + 
    labs(y=expression(paste("Comfimed Cases")))+
    labs(color="Covid19_US_processed") +
    scale_x_date(date_labels = "%Y-%m-%d") 

print(US_State_Trend)
```


### Trend of total confirmed case of top ten states in the U.S. (except NY and NJ)
```{r, results='hide', message=FALSE, warning=FALSE}
TopTen_States <- Covid19_US_processed %>%
  filter(time >= as.Date("2020-04-16")) %>%
  filter(state!= "New York state" & state!= "New Jersey" & state!= "Unpublishedsources") %>%
  top_n(10, cum_confirm) %>%
  arrange(desc(cum_confirm))

TopTen_States <- pull(TopTen_States, state)

TenStates_US <- Covid19_US_processed %>% 
  filter(state %in% TopTen_States) 

str(TenStates_US)

# Save as csv
write.csv(TenStates_US,
          file = "./Data/Processed/TenStates_US (except NY and NJ).csv")
```


```{r, fig.cap=paste("Trend of top ten states in United States (except NY and NJ)"), echo = FALSE}
# Create a ggplot 
TenStates_trend <- ggplot(TenStates_US, aes(x = time, y =cum_confirm, color=state)) +
  geom_line(alpha = 0.95, size = 1) +
  geom_text_repel(aes(label=state),
                      function(TenStates_US) 
                      TenStates_US[TenStates_US$time == as.Date("2020-04-16"),])+
  mytheme +
  theme(legend.position = "none") +
  labs(x=expression(paste("Time"))) + 
  labs(y=expression(paste("Comfimed Cases"))) +
  labs(color="state") +
  scale_x_date(date_labels = "%Y-%m-%d")


print(TenStates_trend)
```

## Analysis trend of dead and heal cases number and rate
### China

```{r, results='hide', message=FALSE, warning=FALSE}
str(Covid19_China_processed)

names(Covid19_China_processed)[1] <- "Province"
names(Covid19_China_processed)[4] <- "Dead Rate"
names(Covid19_China_processed)[6] <- "Heal Rate"

ChinaTrend_gather <- tidyr::gather(Covid19_China_processed, "Type","Rate", 4,6)
head(ChinaTrend_gather)

```


```{r, fig.cap=paste("Compare total death rate and heal rate in each province (China)")}
China_death_heal_Trend <- ggplot(ChinaTrend_gather, aes(fill=Type, y=Rate, x=Province)) + 
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(x=expression(paste("Province Name"))) + 
  labs(y=expression(paste("Rate"))) +
  labs(fill = "Rate Types") +
  mytheme + 
  theme(legend.position = "right") +
  labs(title="Compare death rate and heal rate in China province") 
print(China_death_heal_Trend)
```

**Trend of confirms, deaths,and heal number in China**

```{r, results='hide', message=FALSE, warning=FALSE}
# Change Historical_China_processed dataset's Column names
str(Historical_China_processed)

Historical_China_processed2 <- Historical_China_processed %>%
  group_by(time,country) %>%
  summarise(total_confirm = sum(cum_confirm),
            total_heal = sum(cum_heal),
            total_dead = sum(cum_dead))
head(Historical_China_processed2)

names(Historical_China_processed2)[1] <- "Time"
names(Historical_China_processed2)[3] <- "Total Confirm"
names(Historical_China_processed2)[4] <- "Total Heal"
names(Historical_China_processed2)[5] <- "Total Dead"

Historical_China_gather <- tidyr::gather(Historical_China_processed2, "Type", "Number", 3:5)
str(Historical_China_gather)
```

```{r, fig.cap=paste("Trend of deaths, confirms, and heal number in China")}
Historical_China_gather.plot <- ggplot(Historical_China_gather, aes(Time, Number, color = Type)) +
  geom_point() + 
  geom_line() + 
  labs(x=expression(paste("Time"))) + 
  labs(y=expression(paste("Total number")))+
  mytheme + 
  theme(legend.position = "right") +
  scale_x_date(date_labels = "%Y-%m-%d") +
  labs(title="Trend of deaths, confirms, and heal numberin China") 
print(Historical_China_gather.plot)
```

**Trend of confirms, deaths,and heal number in China (12/1/2019-2/15/2020")**

```{r, fig.cap=paste("Trend of deaths, confirms, and heal number in China (12/1/2019-2/15/2019")}
Historical_China_limt.plot <- ggplot(Historical_China_gather, aes(Time, Number, color = Type)) +
  geom_point() + 
  geom_line() + 
  labs(x=expression(paste("Time"))) + 
  labs(y=expression(paste("Total number")))+
  mytheme + 
  theme(legend.position = "right") +
  scale_x_date(date_labels = "%Y-%m-%d", 
    limits = c(as.Date("2019-12-1"), as.Date("2020-02-15"))) +
  ylim(c(0, 500)) +
  labs(title="Trend of deaths, confirms, and heal numberin China") 
print(Historical_China_limt.plot)
```

**Trend of confirms, deaths,and heal number top ten provines, China** 

```{r, results='hide', message=FALSE, warning=FALSE}
head(TenProvinces_China)

names(TenProvinces_China)[1] <- "Time"
names(TenProvinces_China)[3] <- "Total Confirm"
names(TenProvinces_China)[4] <- "Total Heal"
names(TenProvinces_China)[5] <- "Total Dead"

TenProvinces_China_gather <- tidyr::gather(TenProvinces_China, "Type", "Number", 3:5)
str(TenProvinces_China_gather)
```


```{r, fig.cap=paste("Trend of deaths, confirms, and heal numberin top ten provine, China (expect Hubei)")}
TenProvinces_China_gather.plot <-
  ggplot(TenProvinces_China_gather, aes(x = Time, y = Number, color = Type)) +
  geom_line() +
  geom_point() +
  labs(x=expression(paste("Time"))) + 
  labs(y=expression(paste("Total number")))+
  mytheme + 
  theme(legend.position = "right") +
  scale_x_date(date_labels = "%Y-%m-%d", 
    limits = c(as.Date("2020-1-15"), as.Date("2020-02-15"))) +
  ylim(c(0, 500)) +
  facet_wrap(.~province)
print(TenProvinces_China_gather.plot)

```

### Hubei province, China

**Trend of confirms, deaths,and heal number in top five cities, Hubei(12/1/2019-3/15/2020)**

```{r, results='hide', message=FALSE, warning=FALSE}
head(Historical_Hubei)

# Select top five cities
TopFive_Hubei <- Historical_Hubei %>%
  filter(time >= as.Date("2020-04-16")) %>%
  top_n(5, cum_confirm) %>%
  arrange(desc(cum_confirm))

TopFive_Hubei <- pull(TopFive_Hubei, city)

TopFiveCity_Hubei <- filter(Historical_Hubei, city %in% TopFive_Hubei) %>%
  arrange(desc(cum_confirm))

head(TopFiveCity_Hubei)

# Change column names
names(TopFiveCity_Hubei)[1] <- "Time"
names(TopFiveCity_Hubei)[5] <- "Total Confirm"
names(TopFiveCity_Hubei)[6] <- "Total Heal"
names(TopFiveCity_Hubei)[7] <- "Total Dead"

Historical_Hubei_gather <- tidyr::gather(TopFiveCity_Hubei, "Type", "Number", 5:7)
head(Historical_Hubei_gather)
```


```{r, fig.cap=paste("Trend of confirms, deaths,and heal number in top five cities, Hubei(12/1/2019-3/15/2020)")}
Historical_Hubei_gather.plot <-
  ggplot(Historical_Hubei_gather, aes(x = Time, y = Number, color = Type)) +
  geom_line() +
  geom_point() +
  labs(x=expression(paste("Time"))) + 
  labs(y=expression(paste("Total number")))+
  mytheme + 
  theme(legend.position = "right") +
  scale_x_date(date_labels = "%Y-%m-%d", 
    limits = c(as.Date("2019-12-1"), as.Date("2020-03-15"))) +
  ylim(c(0, 5000)) +
  facet_wrap(.~city)
print(Historical_Hubei_gather.plot)

```



### United States

```{r, results='hide', message=FALSE, warning=FALSE}
# Change Historical_USTrend dataset's Column names
str(Historical_USTrend)

Historical_USTrend2 <- Historical_USTrend
names(Historical_USTrend2)[1] <- "Time"
names(Historical_USTrend2)[3] <- "Total Confirm"
names(Historical_USTrend2)[4] <- "Total Heal"
names(Historical_USTrend2)[5] <- "Total Dead"

USTrend_gather <- tidyr::gather(Historical_USTrend2, "Type", "Number", 3:5)
str(USTrend_gather)
```


```{r, fig.cap=paste("Trend of confirms, deaths,and heal number in the U.S."), echo = FALSE}
US_Trend <- ggplot(USTrend_gather, aes(Time, Number, color = Type)) +
  geom_point() + 
  geom_line() + 
  labs(x=expression(paste("Time"))) + 
  labs(y=expression(paste("Total number")))+
  mytheme + 
  theme(legend.position = "right") +
  scale_x_date(date_labels = "%Y-%m-%d", 
    limits = c(as.Date("2020-01-15"), as.Date("2020-04-16"))) +
  labs(title="Trend of deaths, confirms, and heal number in US") 
print(US_Trend)


```


**Heal and dead plot for top ten states**

```{r, results='hide', message=FALSE, warning=FALSE}
head(TenStates_US)

names(TenStates_US)[1] <- "Time"
names(TenStates_US)[4] <- "Total Confirm"
names(TenStates_US)[5] <- "Total Heal"
names(TenStates_US)[6] <- "Total Dead"

TenStates_US_gather <- tidyr::gather(TenStates_US, "Type", "Number", 4:6)
str(TenStates_US_gather)
```


```{r, fig.cap=paste("Trend of confirms, deaths,and heal number in top ten states"), echo = FALSE}
US_Topten_Trend <-
  ggplot(TenStates_US_gather, aes(x = Time, y = Number, color = Type)) +
  geom_line() +
  geom_point() +
  labs(x=expression(paste("Time"))) + 
  labs(y=expression(paste("Total number")))+
  mytheme + 
  theme(legend.position = "right") +
  facet_wrap(.~state)
print(US_Topten_Trend)

```



\newpage

# Summary and Conclusions

This Study uses visualization way to depict incresed confirmed cases trend on China, United States, and global level, and find their patterns.


According to the information I have obtained from the news, the epidemic situation of Russia ’s COVID-19 has deteriorated rapidly in the past few days (mid-April). The number of cases of COVID-19 detected in Helongjiang from Russia has increased, which makes it the most “outside import” case in China.

According to the news I saw, the reason is that the new cases in early March were mainly from Hong Kong people who participated in two tours to India and Egypt. The new cases increased rapidly in the second half, mainly from people returning from overseas International students and Hong Kong people who have settled in foreign countries, as well as the Languifang Bar infection group.


In the furture research, I will explore the relationship of Covid-19 cases number with high-risk susceptible population and race. Besides, I would like to explore the relationship of Covid-19 with social, economic fields, such as states GDP and states crime rate.


\newpage

# References

Tianzhi Wu, Erqiang Hu, Xijin Ge*, Guangchuang Yu*. Open-source analytics tools for studying the COVID-19 coronavirus outbreak. medRxiv, 2020.02.25.20027433. doi: https://doi.org/10.1101/2020.02.25.20027433*
